---
title: "Analysis of Goodreads books"
output: html_notebook
---

# load data

```{r}
library(tidyverse)
books <- read_csv("data/books.csv", 
                  quote = "") 
```

# review of dataset

```{r}
dim(books)
```

```{r}
names(books)
```

```{r}
head(books, 5)
```

```{r}
tail(books, 5)
```

```{r}
glimpse(books)
```

# clean headings
```{r}
books <- books %>%
  janitor::clean_names()
```

```{r}
glimpse(books)
```

# check for missing data

```{r}
count_na <- function(col){
  sum(is.na(col))
}
books %>%
  summarise(across(everything(), count_na))
```

#look at the rows with NAs

```{r}
books %>% 
  filter(is.na(ratings_count))

```

# remove missing values

```{r}
books_clean <- books %>%
  drop_na()
```

# reduce no of columns to meaningful set

```{r}
books_subset <- books_clean %>%
  select(title, authors, publisher, num_pages, average_rating, ratings_count, text_reviews_count)

books_subset
```


# look for "NOT A BOOK" in author

```{r}
books_subset %>%
  filter(authors == "NOT A BOOK")
```

# add category for books/audio

```{r}
books_categorised <- books_subset %>%
  mutate(category = case_when(
    authors == "NOT A BOOK" ~ "audio",
    TRUE ~ "book"
  ))
```

```{r}
books_only <- books_categorised %>%
  filter(category == "book")

books_only %>%
  distinct(category)
```
# get rid of books with no rating

```{r}
books_with_ratings <- books_only %>%
  mutate(ratings_count = na_if(ratings_count, 0)) %>%
drop_na()

books_with_ratings
```



# Review of Books

# mean of ratings

```{r}
mean_rating <- books_with_ratings %>%
  summarise(mean = mean(average_rating))

mean_rating
```

# median of rating

```{r}
median_rating <- books_with_ratings %>%
  summarise( median = median(average_rating))

median_rating
```

# top 10 books

```{r}
  books_with_ratings%>%
  group_by(category) %>%
  slice_max(average_rating, n = 10) %>%
  arrange(average_rating)
```

# bottom 10 books

```{r}
 books_with_ratings %>%
  select(title, average_rating) %>%
  slice_min(average_rating, n = 10) %>%
  arrange(desc(average_rating))

```


# book reviews above average rating

```{r}
reviews_above_avg <- books_with_ratings %>%
  mutate(difference_frm_avg = average_rating - mean(average_rating)) %>%
  filter(difference_frm_avg > 0)

reviews_above_avg
```


# Review of publishers

# no of publishers

```{r}
books_with_ratings %>%
distinct(publisher)  
```


# mean by pulisher

```{r}
publisher_mean <- books_with_ratings %>%
  group_by(publisher) %>%
  summarise( mean = mean(average_rating))

publisher_mean
```


# top 10 publisher

```{r}
publisher_mean %>%
  slice_max(mean, n = 10) %>%
  arrange(mean)

```

 # bottom 10 publishers
 
```{r}
publisher_mean %>%
  slice_min(mean, n = 10) %>%
  arrange(desc(mean))
```
 
 
 # review by author
 
```{r}
books_with_ratings %>%
distinct(authors)  
```

# mean by author

```{r}
authors_mean <- books_with_ratings %>%
  group_by(authors) %>%
  summarise( mean = mean(average_rating))

authors_mean

```

# top 10 authors

```{r}
authors_mean %>%
  slice_max(mean, n = 10, with_ties = FALSE) %>%
  arrange(mean)

  
```

# bottom 10 authors

```{r}
authors_mean %>%
  slice_min(mean, n = 10) %>%
  arrange(mean)
```



# categorise books by number of pages

```{r}
books_with_ratings %>%
  summarise(range(num_pages))
```


```{r}
book_size <- books_with_ratings %>%
  mutate(book_size = case_when(
    num_pages >1000 ~ "large",
    num_pages >100 ~ "medium",
    TRUE ~ "small")
    )

book_size
```


# ratings per book size

```{r}
book_size %>%
  group_by(book_size) %>%
  summarise(mean = mean(average_rating)) %>% 
  arrange(desc(mean))
```

 # categorise by rating
 
```{r}
books_with_ratings %>%
  summarise(range(average_rating))
```
 
```{r}
rating_category <- books_with_ratings %>%
  mutate(recommendation = case_when(
    average_rating > 3 ~ "highly recommended",
    average_rating > 2 ~ "recommended",
    TRUE ~ "not recommended"
  ))

rating_category
```
 
 
 # review of Harry Potter books
 
 # group by Harry Potter books
 
```{r}
harry_potter_books <- 
filter(books_with_ratings, grepl('Harry Potter', title)
)

harry_potter_books
```

# No of reviews

```{r}
harry_potter_review_count <- harry_potter_books %>%
  group_by(title) %>%
  summarise(ratings_count = sum(ratings_count)) %>%
  arrange(desc(ratings_count))

harry_potter_review_count
```


# Top most rated Harry Potter books

```{r}
harry_potter_review_count %>%
  head(ratings_count, n = 5)
```


# highest rated Harry Potter books

```{r}
harry_potter_highest_ratings <- harry_potter_books %>%
  select(title, average_rating) %>%
  slice_max(average_rating, n = 5, with_ties = FALSE)
  

harry_potter_highest_ratings
```


# lowest rated Harry Potter books

```{r}
harry_potter_highest_ratings <- harry_potter_books %>%
  select(title, average_rating) %>%
  slice_min(average_rating, n = 5, with_ties = FALSE)
  

harry_potter_highest_ratings
```