---
title: "Introduction to `ggplot2` homework"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
    pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

# Homework Info

This homework recaps what you learnt in class today. 

You'll be looking at the `backpack` dataset contained within the `CodeClanData` package using tidyverse. You'll need to load the appropriate packages first.In all plots you make, ensure your labels are appropriate 

**You must write down a short summary of what each plot tells you in order to meet MVP.** Being an analyst is not just the ability to write code, you need to be able to interpret the visualisations and translate the relationships found (or not found) into plain language to explain to non-technical decision makers. Getting as much practice of this as you can while on this course is just (if not more!) important than the code!



# MVP

<br>

**Question 1.**  

Load in the libraries you need, and look at the `backpack` data. Take note of variable names and dimensions.   

```{r}
library(ggplot2)
library(tidyverse)
library(CodeClanData)

head(`backpack`)

```



**Question 2.**

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution. 
```{r}
ggplot(backpack) +
  geom_bar(aes(x = backpack_weight)) +
  ylab("number of students") +
  xlab("backpack weight") +
  labs(title = "Distribution of backback weight") + 
  theme_minimal()

# this graph shows the how many students are carrying each backpack weight. It shows most students carry weights of between 5kg and 15kg. With the most common weight being 10kg.
```


**Question 3.**  

Are there differences in backpack weight between sex?   

```{r}
ggplot(backpack) +
  geom_bar(aes(x = backpack_weight, fill = sex)) +
  ylab("number of students") +
  xlab("backpack weight") +
  labs(title = "Distribution of backback weight") + 
  theme_minimal() +
  facet_grid(~ sex)

# this graph shows how each weight is split between male and female. You can see from the female graph the a higher number of females back weight is more concentrated on certain weights where the males is more evenly split over all of the weights
```
**Question 4.**  

What is the relationship between body weight and backpack weight?

```{r}
ggplot(backpack) +
  aes(x = body_weight, y = backpack_weight, group = backpack_weight) +
  geom_point()

# this shows that there is no real correlation between body weight and backback weight given the data provided
```



**Question 5.**  

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.  

```{r}
ggplot(backpack) +
  aes(x = body_weight, y = backpack_weight, group = backpack_weight, colour = sex) +
  geom_point() +
  facet_grid(~ sex)

# these graphs show that the body weight of males is higher but no significant increase in the backpack weight they carry
```

**Question 6.**  

What **proportion** of students report back pain?   

```{r}
back_problems <- backpack %>% 
  mutate(back_problems = as.logical(back_problems)) 
ggplot(back_problems) +
  aes(x = back_problems, fill = back_problems) +
  geom_bar() +
  labs(
    x = "Suffering from back pain",
    y = "Percentage of students"
  )

# This graph shows that 32% of students suffer from back pain
```



**Question 7.**  

Is this proportion different between men and women? 
```{r}
back_problems <- backpack %>% 
  mutate(back_problems = as.logical(back_problems)) 
ggplot(back_problems) +
  aes(x = back_problems, fill = back_problems) +
  geom_bar() +
  labs(
    x = "Suffering from back pain",
    y = "Percentage of students"
  ) +
  facet_grid(~ sex)

 # These graphs show that there is a much higher cahnce of back pain if you are a female
```



**Question 8.**  

Now you've investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both `backpack_weight` and `body_weight` as well as the `ratio`.   
```{r}
ggplot(backpack) +
  geom_col( aes(x= body_weight, y = backpack_weight)) +
  facet_grid(back_problems ~ sex)

# these graphs show that the females that have back problems usually weigh less (<150) and while the backpack weigh can vary it usually weighs over 20. For males all back problems occur between 150 and 200 with a backpack weight of between 30 and 40.
```

*Hint: faceting would be useful here*.   

# Extensions

<br>

**Question 1.** 

Does reported backpain differ across the year groups? Ignore year 0 in your data.  


**Question 2.** 

Is there a difference between how heavy the bags those in different years carry?


