---
title: "Clustering homework"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


* You have been provided some customer data from a shopping centre. Do some exploratory analysis on the data. Comment on findings. 

```{r}
library(tidyverse)
library(janitor)
library(broom)
library(factoextra)

customers <- read_csv("data/mall_customers.csv") %>% 
  clean_names() %>% 
  select(-c(customer_id))

head(customers)
```
```{r}
library(GGally)

ggpairs(customers)
```

```{r}
ggplot(customers) +
  geom_col(aes(x = gender, y = spending_score_1_100)) 
```
```{r}
ggplot(customers) +
  geom_point(aes(x = annual_income_k, y = spending_score_1_100)) 
```

From the analysis there you can see males spend more than females but there does not seem to be any correaltion between annual income and spending score.

* We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers. 


* Perform k-means clustering and chose a value of k.

# select the two columns and scale the data
```{r}
cust_scale <- customers %>%
              select(annual_income_k, spending_score_1_100) %>% 
                mutate_if(is.numeric, scale)

cust_scale
```

#calculate our clusters, tidy up the data, augment the clusters on to the original data, and use the glance function to see what was contained within our model.

```{r}
max_k <- 20 

k_clusters <- tibble(k = 1:max_k) %>%
  mutate(
    kclust = map(k, ~ kmeans(cust_scale, .x, nstart = 25)), 
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, customers)
  )

k_clusters
```
# unnest the data
```{r}
clusterings <- k_clusters %>%
  unnest(glanced)

clusterings
```

# perform the elbow test
```{r}
ggplot(clusterings, aes(x=k, y=tot.withinss)) +
  geom_point() +
    geom_line() +
    scale_x_continuous(breaks = seq(1, 20, by = 1))
```
This looks like the elbow is ar 3 and we should have a k of 3

# look at the silhouette coefficience
```{r}
fviz_nbclust(cust_scale, 
             kmeans, 
             method = "silhouette", 
             nstart = 25)
```
here this indicates a k = 5 should be used 

# perform a gap analysis

```{r}
fviz_nbclust(cust_scale, 
             kmeans, 
             method = "gap_stat", 
             nstart = 25, 
             k.max = 10)
```
Again this indicates a k = 5 should be used and has a gap statistic over 0.5 which is a good indicator

Overall the indicators point to k of 5 being the best number of clusters


* Visualise the clustering for your chosen value of k.

```{r}
 clusterings %>% 
  unnest(cols = c(augmented)) %>%
  filter(k <= 5) %>%
  ggplot(aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point(aes(color = .cluster))
```



* Do you think the clustering seems a good fit for this data?

From the visualisation this looks like the data for these two variables is a good fit for the data.

* Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the 'Segmentation & clustering intro' lesson). 

Segment 5 have a low annual income and a high spending score
Segment 5 have a high annual income and a high spending score




