---
title: "Text Mining Homework"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE, fig.align = "center", message = FALSE, warning = FALSE)
```

# MVP

Using the dataset `austen_books()` from the package `janeaustenr`:

```{r}
library(tidytext)
library(janeaustenr)
library(tidyverse)
```


1. Find the most common words in both Pride & Prejudice and Sense & Sensibility.

```{r}
titles <- c("Pride and Prejudice", "Sense and Sensibility")

books <- list(prideprejudice, sensesensibility)

books <- purrr::map_chr(books, paste, collapse = " ") %>% 
  paste(collapse = " ")

```


```{r}
all_books_df <- tibble(
  title = titles,
  text = books
) %>%
  unnest_tokens(word, text)

head(all_books_df)
```

```{r}
compare <- tibble(
  titles = c("Pride and Prejudice", "Sense and Sensibility"),
  text = books
)
```

```{r}
compare %>%
  unnest_tokens(word, text) %>%
  count(titles, word) %>%
  bind_tf_idf(word, titles, n) %>%
  arrange(titles, desc(n)) %>%
  group_by(titles)
```


2. Find the most common words in both Pride & Prejudice and Sense & Sensibility, not including stop words.

```{r}
compare %>%
  unnest_tokens(word, text) %>%
  count(titles, word) %>%
  bind_tf_idf(word, titles, n) %>%
  arrange(titles, desc(n)) %>%
  group_by(titles) %>% 
  anti_join(stop_words) 
```


3. Find the most common sentiment words in both Pride & Prejudice and Sense & Sensibility.

```{r}
all_books_df %>%
  inner_join(get_sentiments("bing")) %>% 
  count(word, sort = TRUE)
```


# Extension

Taking your results above. Can you create a plot which visualises the differences between the books?  
